### 2013-05-07 doc/icml2013/abstract.tex

Discriminative latent-variable models are typically learned using
EM or gradient-based optimization, which suffer from local optima.
In this paper, we develop a new computationally efficient
and provably consistent estimator for the mixture of linear regressions,
a simple instance of discriminative latent-variable models.
Our approach relies on a low-rank linear regression to recover
a symmetric tensor, which can be factorized into the parameters
using the tensor power method.
We prove rates of convergence for our estimator
and provide an empirical evaluation illustrating
its strengths relative to local optimization (EM).

>>>>

Discriminative latent-variable models are typically learned using
EM or gradient-based optimization, which suffer from local optima.
In this paper, we develop a new computationally efficient
and provably consistent estimator for a mixture of linear regressions,
a simple instance of discriminative latent-variable models.
Our approach relies on a low-rank linear regression to recover
a symmetric tensor, which can be factorized into the parameters
using the tensor power method.
We prove rates of convergence for our estimator
and provide an empirical evaluation illustrating
its strengths relative to local optimization (EM).

### 2013-05-07 doc/icml2013/evaluation.tex

>>>>

In the previous section, we showed that Spectral Experts is a consistent
algorithm that converges to the global solution at a $\frac{1}{\sqrt}$
rate. In this section, we will support these theoretical results with
experiments on simulated data. We will see that in the finite sample
regime, even when Spectral Experts does not converge to the optimal
answer, it provides robust initialization for local methods like EM.

### 2013-05-07 doc/icml2013/model.tex

The mixture of linear regressions model \citep{VieleTong2002} defines
a conditional distribution over a response $y \in \Re$
given covariates $x \in \Re^d$.
The generation of $y$ given $x$ involves two steps:
(i) draw the mixture component $h \in [K]$ according to mixture proportions
$\pi = (\pi_1, \dots, \pi_K)$;
and (ii) generate $y$ from a Gaussian centered at $\beta_h^\top x$ with
variance $\sigma^2$.
\begin{eqnarray*}
  h &\sim& \mult(\pi), \\
  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 }.
\end{eqnarray*}
The parameters of the model are $\theta = (\pi, B)$,
where $\pi \in \Re^d$ are the mixture proportions and
$B = [\beta_1 \mid \dots \mid \beta_K] \in \Re^{d \times K}$
are the regression coefficients.
We assume the variance of the observation noise $\sigma^2$ is known and that the mixture components are independent.
The learning problem is stated as follows:
given $n$ i.i.d. samples $(\xn{1}, \yn{1}), \dots, (\xn{n}, \yn{n})$
drawn from the model with some unknown parameters $\theta^*$,
return an estimate of the parameters $\hat\theta$.

% History of model
The mixture of linear regressions model has been applied
in the statistics literature for modeling music perception, where $x$ is the
actual tone and $y$ is the tone perceived by a musician \cite{VieleTong2002}.
% Reference mixture of experts.
% Acknowledge that mixture proportions % can't depend on $x$
The model is an instance of hierarchical mixture of experts
\cite{jacobs91experts}, in which the mixture proportions are allowed to depend
on $x$, known as a gating function.
This dependence allow the experts to be localized in input space,
providing more flexibility, but we do not consider this dependence in our model.

% Typically people use EM, but the maximum marginal likelihood is
% non-convex.
The estimation problem for mixture of linear regressions is difficult because
the mixture components $h$ are unobserved,
resulting in a non-convex log marginal likelihood.
The parameters are typically learned using
expectation maximization (EM) or Gibbs sampling \cite{VieleTong2002},
which suffers from local optima.
In the next section, we present a new algorithm
that sidesteps the local optima problem entirely.
%(we show this empirically in \sectionref{sec:evaluation}).
%We provide several instances where this leads
%to poor local optima in \sectionref{sec:evaluation}.

>>>>

The mixture of linear regressions model \citep{VieleTong2002} defines
a conditional distribution over a response $y \in \Re$
given covariates $x \in \Re^d$.
The generation of $y$ given $x$ involves two steps:
(i) draw the mixture component $h \in [k]$ according to mixture proportions
$\pi = (\pi_1, \dots, \pi_k)$;
and (ii) generate $y$ from a Gaussian centered at $\beta_h^\top x$ with
variance $\sigma^2$.
\begin{eqnarray*}
  h &\sim& \mult(\pi), \\
  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 }.
\end{eqnarray*}
The parameters of the model are $\theta = (\pi, B)$,
where $\pi \in \Re^d$ are the mixture proportions and
$B = [\beta_1 \mid \dots \mid \beta_k] \in \Re^{d \times k}$
are the regression coefficients.
We assume the variance of the observation noise $\sigma^2$ is known and that the mixture components are independent.
The learning problem is stated as follows:
given $n$ i.i.d. samples $(\xn{1}, \yn{1}), \dots, (\xn{n}, \yn{n})$
drawn from the model with some unknown parameters $\theta^*$,
return an estimate of the parameters $\hat\theta$.

% History of model
The mixture of linear regressions model has been applied
in the statistics literature for modeling music perception, where $x$ is the
actual tone and $y$ is the tone perceived by a musician \cite{VieleTong2002}.
% Reference mixture of experts.
% Acknowledge that mixture proportions % can't depend on $x$
The model is an instance of the hierarchical mixture of experts
\cite{jacobs91experts}, in which the mixture proportions are allowed to depend
on $x$, known as a gating function.
This dependence allow the experts to be localized in input space,
providing more flexibility, but we do not consider this dependence in our model.

% Typically people use EM, but the maximum marginal likelihood is
% non-convex.
The estimation problem for a mixture of linear regressions is difficult because
the mixture components $h$ are unobserved,
resulting in a non-convex log marginal likelihood.
The parameters are typically learned using
expectation maximization (EM) or Gibbs sampling \cite{VieleTong2002},
which suffers from local optima.
In the next section, we present a new algorithm
that sidesteps the local optima problem entirely.
%(we show this empirically in \sectionref{sec:evaluation}).
%We provide several instances where this leads
%to poor local optima in \sectionref{sec:evaluation}.

### 2013-05-07 doc/icml2013/theory.tex

The proof of the theorem has two parts.
First, we bound the error in the compound parameters estimates $\hat M_2,\hat M_3$
using results from \citet{Tomioka2011}.
Then, we use results from \citet{AnandkumarGeHsu2012} to convert this error
into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$
derived from the robust tensor power method.
But first, let us study a more basic property: identifiability.

>>>>

The proof of the theorem has two parts.
First, we bound the error in the compound parameters estimates $\hat M_2,\hat M_3$
using results from \citet{Tomioka2011}.
While the $R^{12}$ term looks formidable, it is unavoidable since we
perform regression on the third moment of the data. Classically, the
number of samples required is norm squared of the covariance matrix,
which itself is bounded by the norm squared of the data, $R^3$. This
third-order dependence also shows up in the rate the regression
coefficients are annealed; the cubic terms bound each of $\epsilon^3$,
$\beta_h^3$ and $\|(x\tp{3})\tp{2}\|_F$ with high probability.

Next, we use results from \citet{AnandkumarGeHsu2012} to convert this error
into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$
derived from the robust tensor power method.
But first, let us study a more basic property: identifiability.
